---
title: "Technical Memo - Template (DRAFT)"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    collapsed: no
    fig_caption: yes
    highlight: espresso 
    number_sections: yes
    smooth_scroll: yes
    theme: sandstone
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(here)

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)

#trying to source .Rmd but both are broken
#source(here("doc/fraser-pinks.Rmd"))
#rmarkdown::render(input = here("doc/fraser-pinks.Rmd"))


```


*The purpose of this template is to document the Precautionary Approach framework components for Pacific Salmon that Science Branch is responsible for producing. This memo is not meant to provide an assessment of stock status. The template is set up so that:*

  - *main body = advice/information for managers who need to determine the reference points they are responsible when implementing the PA/FSP*
  - *appendices = technical information (e.g. additional results) of interest & scientific documentation necessary to reproduce the analysis*

# Summary  

```{r summary-table}
summary_df <- data.frame(attribute = c("SMU",
                                       "biological benchmark",
                                       "advice for joint science",
                                       "other biological benchmarks to consider",
                                       "next year of re-evaluation",
                                       "recommended approach if points are NOT re-evaluated 
                                       within timeframe",
                                       "reccomended threshold to trigger a rebuilding plan"),
                         data = c("Fraser Pink Salmon",
                                  "80% Smsy",
                                  "fish them??",
                                  "Smsy in context of changing demographics (e.g. lower body 
                                  mass, lower fecundity)",
                                  "2025",
                                  "figure it out!",
                                  "in bottom right quadrant of kobe plot"))

kbl(summary_df, 
      col.names = NULL, 
      align = "c") |>
  kable_styling(bootstrap_options = "hover")
```

# Background  
*Suggest some mostly generic language that can be cut & pasted into each PA/FSP memo re: purpose of memo, what it is/is not. PA components that Science is directly responsible for,  jointly responsible for, and provides support to Fisheries Management for.*

*If helpful, could (generically) include what some biological benchmarks are (e.g. 80% Smsy, Umsy, etc.) and why they may be considered / informative when determining management reference points*

include terminology that we want to use for all PA/FSP memos: 

- biological benchmarks
- biological reference points (?)
- management reference points
- etc.

# Biological Context  
- stock aggregation units/structure - i.e. how does CU, MU & SMU relate for this group?
- are there any previous estimates of PA components? if yes: what are they and when were they estimated?

# Harvest Control Rule  
- description / diagram of current HCR
- ID any "implied" management values that come from the HCR (will be useful for comparative purposes to biological benchmarks)

FR PK doc: 

- ***ADD HCR plots from ppt (Aug2022)***

# Data summary & plots  
- descriptive plots of data used in analysis to show quantity of data available
- any description/caveats re: data quality would be useful to include in this section, as well
- data tables go in appendix

FR PK doc:
```{}
#how tf do I read a plotly fig from one .Rmd into another?
```
- raw escapement & harvest boxplots (3.1)
- table of escapement assessment methods (3.1)

# Methods - summary  
*short, general audience description of methods and how uncertainty was incorporated (or not). Put details that are required for scientific documentation in appendix*

# Results for PA/FSP implementation 
*focus in this section is on results that determine the reference points that Science Branch is responsible / jointly responsible for and biological reference points that managers may want to consider when making decisions re: implementation of PA/FSP reference points*

## Part A: benchmark/reference point values
*table(s) of benchmark values and uncertainty range: for both those that Science is responsible / co-responsible for as well as benchmarks that could help FAM determine management reference points* 

FR PK doc:

- table of benchmarks & CIs (3.3)

## Part B: risk context
*things that may be helpful for managers as they determine management reference points. e.g., kobe plots, comparison of biological benchmarks to: historical data (spawners, harvest rates, etc.), harvest control rules, etc.), risk tables/plots/profiles, sensitivity analyses...*

FR PK doc:

- Sgen & 80%Smsy vs SPN (3.3)
- HR vs Umsy (3.3)
- expected yield & recruitment profiles (3.4)
- Kobe plot (4)

# Sources of Uncertainty  / Time-Varying Considerations
*Q: do we need "sources of uncertainty" here, given above section re: risk context? OR maybe change this explicitly to a section talking about whether time-varying benchmarks should be given consideration? Climate change considerations, etc.*

include in this section:

* recommended timeframe for next evaluation of reference points & benchmarks
* rationale for that timeframe
* recommendation for what to use if re-evaluation does not take place within the recommended timeframe & why
* recommendation for what would trigger a rebuilding plan: how many year(s) of what units/metric (individual years vs running average vs ...) being below LRP would trigger a rebuilding plan, and rationale for recommendation 


FR PK doc (time-varying):

- SR plot, color coded to time (3.2)
- recruitment resids (3.2)


# Other considerations
*Are there other things that Fisheries Management should consider or should know about the results in this memo when developing management benchmarks?*

# Data Sources and Acknowledgements
*Data collection, quality control, management, etc. that the analysis relied upon. Include dates (year) when data set used was received and/or last updated.*

# Appendices
*technical information that is not directly needed for management purposes, but provides additional information about the results, is necessary for documentation of the analysis and/or to allow for future reproducibility*

1. data
2. methods
3. results that are not included in main body of memo - anticipated that these will mostly be diagnostic in nature?
4. sources of uncertainty (assuming that it will be removed from main body of document)
5. references
 