---
title: "Technical Memo - Template (DRAFT)"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    collapsed: no
    fig_caption: yes
    highlight: espresso 
    number_sections: yes
    smooth_scroll: yes
    theme: sandstone
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(rstan)
library(tidyverse)
library(kableExtra)
library(gsl)
library(ggpubr)
library(viridis)
library(plotly)
library(here)

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      dpi=300,
                      fig.align = "center")

options(scipen=1, digits=4)

# Read in data and posterior samples
data <- read.csv(here("data/fr_pk_spw_har.csv"))
stan.fit <- readRDS(here("analysis/output/SS-SR_AR1.stan.fit.rds"))
stan.data <- readRDS(here("analysis/output/SS-SR_AR1.stan.data.rds"))
model_pars <- rstan::extract(stan.fit)
```


*The purpose of this template is to document the Precautionary Approach framework components for Pacific Salmon that Science Branch is responsible for producing. This memo is not meant to provide an assessment of stock status. The template is set up so that:*

  - *main body = advice/information for managers who need to determine the reference points they are responsible when implementing the PA/FSP*
  - *appendices = technical information (e.g. additional results) of interest & scientific documentation necessary to reproduce the analysis*

# Summary  

```{r summary-table}
summary_df <- data.frame(attribute = c("SMU",
                                       "biological reference point",
                                       "joint science-FAM advice",
                                       "other biological reference points to inform management reference points",
                                       "next year of re-evaluation",
                                       "recommended approach if points are NOT re-evaluated 
                                       within timeframe",
                                       "reccomended threshold to trigger a rebuilding plan"),
                         data = c("Fraser Pink Salmon",
                                  "*(Sgen with CI range)*",
                                  "removal reference: e.g., Umsy; X% Umsy; *(insert values with range)* ",
                                  "values which may be useful for determining management reference points (e.g., Smsy, X%Smsy, percentile based values)",
                                  "*(insert year here)*",
                                  "TBD",
                                  "TBD *(include measurement units & frequency)* "))

kbl(summary_df, 
      col.names = NULL, 
      align = "l", 
    caption = "Example of key summary information for a given stock") |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Background  
*Suggest some mostly generic language that can be cut & pasted into each PA/FSP memo re: purpose of memo, what it is/is not. PA components that Science is directly responsible for,  jointly responsible for, and provides support to Fisheries Management for.*

*If helpful, could (generically) include what some biological refpts are (e.g. 80% Smsy, Umsy, etc.) and why they may be considered / informative when determining management reference points*

include terminology that we want to use for all PA/FSP memos: 

- biological reference points (which can include WSP biological benchmarks)
- management reference points
- *Q: Harvest Decision Rule vs Harvest Control Rule??? (have kept as HCRs in this version, can change if that's the new terminology)*
- etc.

# Biological Context  
- stock aggregation units/structure - i.e. how does CU, MU & SMU relate for this group?
- are there any previous estimates of biological reference points? if yes: what are they and when were they estimated?
- include any relevant ecological info, e.g.:
  - age structure
  - migration, spawn timing
  - distribution
  - etc...incl. any trends

# Harvest Control Rule  
- description / diagram of current HCR
- ID any "implied" management values that come from the HCR (will be useful for comparative purposes to biological benchmarks)
- work with FAM / FAM to provide short summary of ancillary mgmt pieces - e.g., time, area, effort control that works with the HCR 

FR PK doc: 

- ***ADD HCR plots from ppt (Aug2022)***

# Data summary & plots  
- descriptive plots of data used in analysis to show quantity of data available
- any description/caveats re: data quality would be useful to include in this section, as well
- data tables go in appendix

```{r esc har, out.height='80%', out.width = '80%', fig.cap= "Total run for Fraser Pink Salmon during the monitoring period"}
escape_harvest <- gather(data[,1:3], type, count, harvest:spawn, factor_key=TRUE) |>
  mutate(count_million = round(count/1000000, 2))

ggplot(escape_harvest, aes(fill=type, y=count_million, x=year)) + 
  geom_bar(position="stack", stat="identity") +
  scale_fill_brewer(palette="Dark2",name = "Return type") +
  xlab("Year") +
  ylab("Total return (millions of fish)") +
  theme_bw()
```

```{r spawner estimation methods}
df <- data.frame(Years = c("1957-1985 ", "1987-1991", "1993-2001", "2003-2007", "2009-2011", "2013-2021"),
                 Method = c("PSC mark-recapture (system specific)", "DFO mark-recapture (system specific)", "DFO mark-recapture (mainstem)", "Test fishery", "Mission post-season", "Mission in-season"),
                 CV = c("20%", "20%", "20%", "50%", "35%", "35%"))

kable(df, 
      align="l",
      caption = "Sampling stanzas and observation error CVs used in state-space model") |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Methods - summary  
*short, general audience description of methods and how uncertainty was incorporated (or not). Put details that are required for scientific documentation in appendix*

# Results for PA/FSP implementation 
*focus in this section is on results that determine the reference points that Science Branch is responsible / jointly responsible for and biological reference points that managers may want to consider when making decisions re: implementation of PA/FSP reference points*

## Part A: biological reference point values

This will be provided at both SMU & CU level (when they are not a 1:1 relationship)

```{r SR data wrangle}
max_samples <- dim(model_pars$lnalpha)

spwn <- exp(model_pars$lnS)
spwn.quant <- apply(spwn, 2, quantile, probs=c(0.05,0.5,0.95))[,1:31]

rec <-exp(model_pars$lnR)
rec.quant <- apply(rec, 2, quantile, probs=c(0.05,0.5,0.95))[,2:32]

brood_t <- as.data.frame(cbind(data$year[1:31],t(spwn.quant), t(rec.quant)))
colnames(brood_t) <- c("BroodYear","S_lwr","S_med","S_upr","R_lwr","R_med","R_upr")

brood_t <- as.data.frame(brood_t) |>
  round(2)

# SR relationship
spw <- seq(0,max(brood_t[,4]),length.out=100)
SR_pred <- matrix(NA,100,max_samples)

for(i in 1:max_samples){
  r <- sample(seq(1,max_samples),1,replace=T)
  a <- model_pars$lnalpha[r]
  b <- model_pars$beta[r]
  SR_pred[,i] <- (exp(a)*spw*exp(-b*spw))
}

SR_pred <- cbind(spw,t(apply(SR_pred,c(1),quantile,probs=c(0.05,0.5,0.95),na.rm=T)))
colnames(SR_pred) <- c("Spawn", "Rec_lwr","Rec_med","Rec_upr")
SR_pred <- as.data.frame(SR_pred) |>
  round(2)

#create spawner df
spwn <- model_pars$S
spwn.quant <- apply(spwn, 2, quantile, probs=c(0.025,0.25,0.5,0.75,0.975))

spwn_df <- as.data.frame(cbind(data$year, t(spwn.quant)))
colnames(spwn_df) <- c("year","lwr","midlwr","mid","midupper","upper")

#create escapement df
er <- model_pars$U
er.quant <- apply(er, 2, quantile, probs=c(0.025,0.25,0.5,0.75,0.975))

er_df <- as.data.frame(cbind(data$year, t(er.quant)))
colnames(er_df) <- c("year","lwr","midlwr","mid","midupper","upper")

#get yield profiles 
spw <- seq(0,max(brood_t[,4]),length.out=100)
Profiles <- array(NA, dim = c(length(spw),max_samples,9))
for(i in 1:max_samples){
  r <- sample(seq(1,1000),1,replace=T)
  a <- model_pars$lnalpha[r]
  b <- model_pars$beta[r]
  R_pred <- spw*exp(a)*exp(-b*spw)
  R_pred[which(!is.finite(R_pred))] <- 0
  pred_spwn <- spw
  yield <- R_pred-pred_spwn
  S_msy_c <- (1 - lambert_W0(exp(1 - a))) / b
  msr_c <- max(R_pred)
  S_msr_c <- pred_spwn[which.max(R_pred)]
  msy_c <- max(yield)
  
  for(j in 1:length(spw)){
    spw_star <- spw[j]
    R_star <- spw_star*exp(a-b*spw_star)
    if(spw_star==0){R_star<-0; spw_star<-0}
    sus_yield <- R_star - spw_star
    U <- ifelse(sus_yield/R_star < 0, 1, sus_yield/R_star)
    Profiles[j,i,1] <- ifelse(sus_yield > 0.9*msy_c, 1, 0) # probability yield > 90% of MSY
    Profiles[j,i,2] <- ifelse(sus_yield > 0.8*msy_c, 1, 0) # probability yield > 80% of MSY
    Profiles[j,i,3] <- ifelse(sus_yield > 0.7*msy_c, 1, 0) # probability yield > 70% of MSY
    Profiles[j,i,4] <- ifelse(R_star > 0.9*msr_c, 1, 0) # probability recruitment > 90% of MSR
    Profiles[j,i,5] <- ifelse(R_star > 0.8*msr_c, 1, 0) # probability recruitment > 80% of MSR
    Profiles[j,i,6] <- ifelse(R_star > 0.7*msr_c, 1, 0) # probability recruitment > 70% of MSR
    if(sus_yield < (0.9*msy_c) & spw_star < S_msy_c){
      Profiles[j,i,7] <- 1
    } else {
      Profiles[j,i,7] <- 0
      }
    if(sus_yield < (0.8*msy_c) & spw_star < S_msy_c){
      Profiles[j,i,8] <- 1
    } else {
      Profiles[j,i,8] <- 0
      }
    if(sus_yield < (0.7*msy_c) & spw_star < S_msy_c){
      Profiles[j,i,9] <- 1
    } else {
      Profiles[j,i,9] <- 0
      }    
        }
}
profiles_median <- cbind(spw,apply(Profiles,c(1,3),mean, na.rm=T))
colnames(profiles_median) <- c("Spawners","90% MSY","80% MSY","70% MSY", "90% MSR","80% MSR","70% MSR", "90% MSY ", "80% MSY ", "70% MSY ")
profiles_median <- as.data.frame(profiles_median)

long_yield_profiles <- profiles_median %>%
  select("Spawners","90% MSY","80% MSY","70% MSY") %>%
  pivot_longer(!Spawners, names_to = "Objective", values_to = "perc")

```

```{r benchmarks}
# Sgen function
get_Sgen <- function(a, b, int_lower, int_upper, SMSY) {
  fun_Sgen <- function(Sgen, a, b, SMSY) {Sgen * a * exp(-b * Sgen) - SMSY}
  Sgen <- uniroot(fun_Sgen, interval=c(int_lower, int_upper), a=a, b=b, SMSY=SMSY)$root
}

# Benchmarks
bench <- matrix(NA,1000,3,
                dimnames = list(seq(1:1000), c("sgen","smsy","umsy")))

for(i in 1:1000){
  r <- sample(seq(1,1000),1,replace=T)
  a <- model_pars$lnalpha[r]
  b <- model_pars$beta[r]
  bench[i,2] <- (1 - lambert_W0(exp(1 - a))) / b # smsy
  bench[i,1] <- get_Sgen(exp(a),b,-1,1/b*2,bench[i,2]) # sgen
  bench[i,3] <- (1 - lambert_W0(exp(1 - a)))# umsy
}
  
bench[,2] <- bench[,2]*0.8
bench.quant <- apply(bench, 2, quantile, probs=c(0.025,0.5,0.975), na.rm=T)

percentiles <- quantile(data$spawn,probs=c(0.25, 0.5))/1000000

benchmarks <- matrix(NA,5,3)
benchmarks[1,] <- c(bench.quant[2,2],bench.quant[1,2],bench.quant[3,2])
benchmarks[2,] <- c(bench.quant[2,1],bench.quant[1,1],bench.quant[3,1])
benchmarks[3,] <- c(bench.quant[2,3],bench.quant[1,3],bench.quant[3,3])
benchmarks[4,1] <- percentiles[1]
benchmarks[5,1] <- percentiles[2]
rownames(benchmarks) <- c("80% Smsy","Sgen","Umsy","25th percentile","50th percentile")
colnames(benchmarks) <- c("median","lower CI","upper CI")

benchmarks.df <- as.data.frame(benchmarks)

kable(data.frame(benchmarks.df), 
      align="l",
      caption = "Table(s) of benchmark values and uncertainty range: for both those that Science is responsible / co-responsible for as well as benchmarks that could help FAM determine management reference points") |> 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Part B: risk context
*things that may be helpful for managers as they determine management reference points. e.g., kobe plots, comparison of biological benchmarks to: historical data (spawners, harvest rates, etc.), harvest control rules, etc.), risk tables/plots/profiles, sensitivity analyses...*

```{r kobe-plot, out.height='80%', out.width = '80%', fig.cap= "Kobe plot showing risk relative to Smsy and Umsy"}
kobe_df <- data.frame(S = spwn_df$mid, 
                      U = er_df$mid, 
                      year = spwn_df$year) |>
  mutate(U_Umsy = round(U/benchmarks[3,1] ,2), 
         S_Smsy = round(S/(benchmarks[1,1]*1.25), 2)) #correct from 80%Smsy to Smsy

p <- ggplot(kobe_df, aes(S_Smsy, U_Umsy)) +
  geom_vline(xintercept = 1, lty = 2) +
  geom_vline(xintercept = 0.8, lty = 3) +
  geom_hline(yintercept = 1, lty = 2) +
  geom_path() +
  geom_point(aes(color = year, text = year), size=3) +
  scale_colour_viridis_c(name="Year") +
  theme_classic() +
  theme(legend.position = "bottom") + #hack to functionally align plots
  labs(y="U/Umsy", x= "S/Smsy")

plotly::ggplotly(p, tooltip = "text") |> 
  config(displayModeBar = FALSE) 
```
  - Q for group: axes --> currently shows using S/Smsy, but one potential mgmt refpt = 80%Smsy (see lighter dotted line); would color coding help? (e.g. red vs green quadrants)
  
```{r Sgen 80 perc Msy, out.width = "75%", fig.cap = "Spawner abundance relative to Smsy with ribbon shading representing 50 and 95% quantiles"}
ggplot(spwn_df, aes(x=year, y = mid), show.legend = FALSE) +
  geom_ribbon(aes(ymin = benchmarks[1,2], ymax = benchmarks[1,3]), fill = rgb(0.5,0.5,0.5, alpha=0.2)) +
  geom_segment(aes(x = 1959 , y = benchmarks[1,1], xend = 2021, yend = benchmarks[1,1], legend = FALSE)) +
  geom_ribbon(aes(ymin = benchmarks[2,2], ymax = benchmarks[2,3]), fill = rgb(0.5,0.5,0.5, alpha=0.2)) +
  geom_segment(aes(x = 1959 , y = benchmarks[2,1], xend = 2021, yend = benchmarks[2,1], legend = FALSE)) +
  geom_line(show.legend = F, color = rgb(1,0,0, alpha=0.2), lwd = 1.5) + 
  geom_ribbon(aes(ymin = lwr, ymax = upper), show.legend = FALSE, fill = rgb(1,0,0, alpha=0.2)) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupper), show.legend = FALSE, fill = rgb(1,0,0, alpha=0.2)) +
  xlab("Year") +
  ylab("Spawner abundance (millions)") +
  annotate("text", label = " 80% Smsy", x = 1962.5, y = 4.90) +
  annotate("text", label = "Sgen", x = 1961, y = 2.05) +
  theme(legend.position = "none") +
  theme_bw()
```

```{r harvest Umsy, out.width = "75%", fig.cap = "Harvest rate and Umsy with ribbon shading representing 50 and 95% quantiles"}
ggplot(er_df, aes(x=year, y = mid), show.legend = F) +
  geom_ribbon(aes(ymin = benchmarks[3,2], ymax = benchmarks[3,3]), fill = rgb(0.5,0.5,0.5, alpha=0.2)) +
  geom_segment(aes(x = 1959 , y = benchmarks[3,1], xend = 2021, yend = benchmarks[3,1], legend = FALSE)) +
  geom_hline(aes(yintercept = benchmarks[3,3]), colour = 'light grey')+
  geom_line(show.legend = F, color = rgb(1,0,0, alpha=0.2), lwd = 1.5) + 
  geom_ribbon(aes(ymin = lwr, ymax = upper), show.legend = F, fill = rgb(1,0,0, alpha=0.2)) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupper), show.legend = F, fill = rgb(1,0,0, alpha=0.2)) +
  coord_cartesian(ylim=c(0,1)) +
  xlab("Year") +
  ylab("Harvest rate") +
  annotate("text", label = "Umsy", x = 2018, y = 0.62) +
  theme(legend.position = "none") +
  geom_abline(intercept = 0, slope = 0,col="dark grey", lty=2)+
  theme_bw()
```

```{r yield and recruit profiles, out.width = "85%", fig.cap = "Yield and recruit profiles"}
a <- ggplot(long_yield_profiles, aes(x = Spawners, y = perc, group = Objective, color = Objective)) +
  geom_line(size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  xlab("Spawners") +
  ylab("Probability") +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank()) +
  theme_bw()  +
  scale_x_continuous(breaks=seq(0,25,5), limits=c(0,25)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()) 
long_rec_profiles <- profiles_median%>%
  select("Spawners","90% MSR","80% MSR","70% MSR")%>%
  pivot_longer(!Spawners, names_to = "Objective", values_to = "perc")
b <- ggplot(long_rec_profiles, aes(x = Spawners, y = perc, group = Objective, color = Objective)) +
  geom_line(size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  xlab("Spawners") +
  ylab("Probability") +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank()) +
  theme_bw() +
  scale_x_continuous(breaks=seq(0,25,5), limits=c(0,25)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())
long_overfish_profiles <- profiles_median%>%
  select("Spawners", "90% MSY ", "80% MSY ", "70% MSY ")%>%
  pivot_longer(!Spawners, names_to = "Objective", values_to = "perc")
c <- ggplot(long_overfish_profiles, aes(x = Spawners , y = perc, group = Objective, color = Objective)) +
  geom_line(size = 1.5) +
  scale_color_viridis(discrete = TRUE) +
  xlab("Spawners (millions)") +
  ylab("Probability ") +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank()) +
  theme_bw() +
  scale_x_continuous(breaks=seq(0,25,5), limits=c(0,25))+
  geom_rug(data = brood_t,
           aes(x = S_med),
           inherit.aes = FALSE,
           sides="b")
g <- ggarrange(a,b,c,nrow =3, labels = c("a", "b", "c"), heights = c(0.75,0.75,0.9))
print(g)
```

# Additional Considerations
*Are there things that Fisheries Management should consider or should know about the results in this memo when developing management benchmarks?*

general things to include in this section:

- sources of uncertainty
- time varying considerations
- changing demographics
- anticipated climate change impacts
- changes in habitat (past & future)
- migration impediments

specific things to include in this section:

* recommended timeframe for next evaluation of reference points & benchmarks
* rationale for that timeframe
* recommendation for what to use if re-evaluation does not take place within the recommended timeframe & why
* recommendation for what would trigger a rebuilding plan: how many year(s) of what units/metric (individual years vs running average vs ...) being below LRP would trigger a rebuilding plan, and rationale for recommendation 

```{r SR plot, out.height='80%', out.width='90%', fig.cap = "Spawner-Recruit relationship. The grey shaded area are the 90% confidince intervals, and the errorbars on points are 95% confidince intervals."}
p <- ggplot() +
  geom_ribbon(data = SR_pred, aes(x = Spawn, ymin = Rec_lwr, ymax = Rec_upr),
              fill = "grey80", alpha=0.5, linetype=2, colour="gray46") +
  geom_line(data = SR_pred, aes(x = Spawn, y = Rec_med), color="black", size = 1) +
  geom_errorbar(data = brood_t, aes(x= S_med, y = R_med, ymin = R_lwr, ymax = R_upr),
                colour="grey", width=0, size=0.3) +
  geom_errorbarh(data = brood_t, aes(x= S_med, y = R_med, xmin = S_lwr, xmax = S_upr),
                 height=0, colour = "grey", height = 0, size = 0.3) +
  geom_point(data = brood_t, 
             aes(x = S_med, 
                 y = R_med, 
                 color=BroodYear, 
                 text= BroodYear, #to pass to hover - can add paste() for context 
                 width = 0.9), 
             size = 3)+
  coord_cartesian(xlim=c(0, 20), ylim=c(0,max(brood_t[,7]))) +
  scale_colour_viridis_c(name = "Brood Year")+
  xlab("Spawners (millions)") +
  ylab("Recruits (millions)") +
  theme_bw() +
  theme(#legend.position = "bottom",  #broken?
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        legend.title = element_text(size=9),
        legend.text = element_text(size=8))+
  geom_abline(intercept = 0, slope = 1,col="dark grey")

plotly::ggplotly(p, tooltip = "text") |>
  config(displayModeBar = FALSE) 
```


```{r recruitment residuals, out.width = "75%", fig.cap = "Recruitment resisuals highlighting years of high/low productivity with ribbon shading representing 50 and 95% quantiles."}
resid <- model_pars$lnresid
resid.quant <- apply(resid, 2, quantile, probs=c(0.025,0.25,0.5,0.75,0.975))[,1:31]

resids <- as.data.frame(cbind(data$year[1:31], t(resid.quant)))
colnames(resids) <- c("year","lwr","midlwr","mid","midupper","upper")

ggplot(resids, aes(x=year, y = mid), show.legend = FALSE) +
  geom_line(show.legend = FALSE, color = rgb(1,0,0, alpha=0.2), lwd = 1.5) + 
  geom_ribbon(aes(ymin = lwr, ymax = upper), show.legend = FALSE, fill = rgb(1,0,0, alpha=0.2)) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupper), show.legend = FALSE, fill = rgb(1,0,0, alpha=0.2)) +
  coord_cartesian(ylim=c(-2,2)) +
  xlab("Brood year") +
  ylab("Recruitment residuals") +
  theme(legend.position = "none") +
  geom_abline(intercept = 0, slope = 0, col = "dark grey", lty = 2) +
  theme_bw()
```


# Data Sources and Acknowledgements
*Data collection, quality control, management, etc. that the analysis relied upon. Include dates (year) when data set used was received and/or last updated.*

# Appendices
*technical information that is not directly needed for management purposes, but provides additional information about the results, is necessary for documentation of the analysis and/or to allow for future reproducibility*

1. data
2. methods
3. results that are not included in main body of memo - anticipated that these will mostly be diagnostic in nature?
4. sources of uncertainty (assuming that it will be removed from main body of document)
5. references
 